---
title: "Lab 14 Homework"
author: "Chelsea Ha"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above. For any included plots, make sure they are clearly labeled. You are free to use any plot type that you feel best communicates the results of your analysis.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(ggmap)
library(leaflet)
``` 

## API Key
```{r}
register_stadiamaps("ece99a72-2ade-4e51-85ff-001ee7169fb5", write = FALSE)
```

## Load the Data
For this homework, we will use the shark attack data to visualize where the attacks occurred. The [data](https://catalog.data.gov/dataset/shark-incident-database-california-56167) are from: State of California- Shark Incident Database.  
```{r message=FALSE, warning=FALSE}
sharks <- read_csv("data/SharkIncidents_1950_2022_220302.csv") %>% 
  clean_names() %>% 
  filter(longitude !="NA" & latitude !="NA") %>% # pulling out NA locations
  mutate(longitude = as.numeric(longitude)) # converting longitude to numeric
```

1. Use the range of the latitude and longitude to build an appropriate bounding box for your map.
```{r}
summary(sharks)
```
```{r}
map1 <- get_stadiamap(bbox = c(left = -125, bottom = 32, right = -114, top = 42), zoom = 6, maptype = "stamen_terrain")
```

- sharks_dups <- sharks %>% 
  distinct(location, .keep_all = TRUE) # remove duplicate locations, but keep the remaining variables
- sharks_dups %>% 
  select(longitude, latitude) %>% 
  summary()  
- lat <- c(32.59, 41.56)
long <- c(-117.1, -124.7)
bbox2 <- make_bbox(long, lat, f=0.3)

2. Load a map from `stamen` in a terrain style projection and display the map.  
```{r}
ggmap(map1)
```
map2 <- get_stadiamap(bbox2, maptype = "stamen_terrain", zoom=7)

3. Build a map that overlays the recorded observations of shark attacks in California.
```{r}
ggmap(map1)+
  geom_point(data = sharks, aes(x = longitude, y = latitude), color = "blue", size = 0.5, alpha = 0.5)+
  labs(title="Shark Attacks in California", x="Longitude", y="Latitude")
  
```


4. Build a map that only shows the fatalities.  
```{r}
ggmap(map1)+
  geom_point(data = sharks %>% filter(injury == "fatal"), 
             aes(x = longitude, y = latitude), 
             color = "red", 
             size = 0.5, 
             alpha = 0.5)+
  labs(title="Fatal Shark Attacks in California", x="Longitude", y="Latitude")
```


5. Build the same map (fatalities only), but this time make it interactive using `leaflet`. Also, color the points red.  
```{r, eval=F, echo=T}
leaflet(sharks %>% filter(injury == "fatal")) %>% 
  addTiles() %>% 
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = 2,
    stroke = FALSE,   
    fillOpacity = 0.7,
    color = "red") %>% 

  fitBounds(lng1=-125, lat1=32, lng2=-114, lat2=42)
```
#Answer to question 5
sharks_fatal %>%
  summarize(
    lon_min = min(longitude, na.rm = TRUE),
    lon_max = max(longitude, na.rm = TRUE),
    lat_min = min(latitude,  na.rm = TRUE),
    lat_max = max(latitude,  na.rm = TRUE))



lon_min <- -123.7783		
lon_max <- -117.28	
lat_min <- 32.855		
lat_max <- 39.57833	



leaflet(sharks_fatal) %>%
  addProviderTiles(providers$Stadia.StamenTerrain) %>%
  addCircleMarkers(
    lng = ~longitude, #use the longitude column for x-coordinates
    lat = ~latitude, #use the latitude column for y-coordinates
    radius = 2, #set marker size (in pixels)
    stroke = FALSE, #remove the outline around each marker
    fillColor="red", #set the fill color of the markers
    fillOpacity = 1 #set the fill opacity of the markers
  ) %>%
  addScaleBar(position = "bottomleft") %>% # add a scale bar to the map
  fitBounds(lng1 = lon_min, lat1 = lat_min, #southwest corner of bounding box
            lng2 = lon_max, lat2 = lat_max) #northeast corner of bounding box





6. Which county has the highest number of recorded shark attacks? Make a map that only shows the attacks in this county. (hint: it may help improve the map if you remove duplicate locations, i.e. multiple attacks at the same location.)
```{r}
sharks %>% 
  group_by(county) %>% 
  summarise(count = n()) %>%
  arrange(desc(count))
```

```{r}
ggmap(map1)+
  geom_point(data = sharks %>% filter(county == "San Diego") %>% distinct(longitude, latitude), 
             aes(x = longitude, y = latitude), 
             color = "blue", 
             size = 0.5, 
             alpha = 0.5)+
  labs(title="Shark Attacks in San Diego County", x="Longitude", y="Latitude")
```


## Knit and Upload
Please knit your work as an .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!